\batcave: Bayesian Analysis Tools for Context-Aware Variant Evaluation

### 2019-09-30 NAR_DRAFT.tex
working on wording for calibration

\begin{figure}
  \begin{center}
  \includegraphics{figures/ppv_wes.pdf}
  \end{center}
  \caption{Calibration of posterior probability for 500X exomes. 
  For realistic thresholds, posterior probabilities are more accurately calibrated for \batcave than MuTect, no matter the concentration of the mutation profile.
  \rngcomment{Need to be careful about what is represented here. Here precision and recall are for all variants \textbf{above} the threshold, correct?}
  }
\label{NAR-ppv_fig}
\end{figure}

In practice, accurate calibration is most important for variants with high posterior probability.
For variants with posterior probability between 60 and 90\%, \batcave is substantially better calibrated than MuTect (Fig.~\ref{NAR-ppv_fig}\&\ref{NAR-ppv_wgs_fig}).
For any posterior probability above 70\%, MuTect has a false positive rate of roughly 8\%, whereas \batcave has a false positive rate that decreases as the posterior probability increases.
The cost of MuTect's compression is recall, which is 3-5\% worse than \batcave at 95\% posterior probability, and 3-7\% worse at 90\%, depending on mutation profile concentration.
At any posterior probability \batcave returns a larger fraction of all positive calls than MuTect. \rngcomment{What does this mean?}
%Figure~\ref{NAR-ppv_wgs_fig} shows results for 100X whole genomes, where the results are similar for precision.

>>>>

\begin{figure}
  \begin{center}
  \includegraphics{figures/ppv_wes.pdf}
  \end{center}
  \caption{Performance of various posterior probability thresholds in 500X exomes. 
  For realistic posterior probability thresholds,  positive predictive values are greater for \batcave than MuTect, no matter the concentration of the mutation profile.
  }
\label{NAR-ppv_fig}
\end{figure}

\bkmcomment{I think there is a reasonable wording here in terms of positive predictive value.
Think of the threshold as the posterior class probability above which we set all probabilities to 1, and below to zero?
Getting better below but I still need to think harder.}
In practice, variant calling is a decision problem, in which a threshold score must be selected above which a variant call is made.
The proper threshold is ideally a choice the user makes based on their particular needs, which often balance precision with recall.
\batcave provides posterior probabilities averaged not only across the whole range of possible probabilities, but in particular at higher probability ranges applicable in most situations.
For posterior probability thresholds between 60 and 90\%, \batcave exhibits substantially better average calibration than MuTect (Fig.~\ref{NAR-ppv_fig}\&\ref{NAR-ppv_wgs_fig}).
As a result, \batcave posterior probabilities provide a useful way to select tumor-specific thresholds. 
For any posterior probability above 70\%, MuTect has a false positive rate of roughly 8\%, whereas \batcave has a false positive rate that decreases as the posterior probability increases.
%The cost of MuTect's compression is recall, which is 3-5\% worse than \batcave at 95\% posterior probability, and 3-7\% worse at 90\%, depending on mutation profile concentration.
At any posterior probability threshold \batcave has recall better than MuTect.
%Figure~\ref{NAR-ppv_wgs_fig} shows results for 100X whole genomes, where the results are similar for precision.
